Probabilistic Inference of Basic Stellar Parameters:\\ Application to Flickering Stars \altaffilmark{\titledag}

### 2016-01-27 version1.0/manuscript.tex
small changes from referee and equation corrections

An alternative approach for inferring model parameters is to allow
relationships between observables to be stochastic.
Unlike the deterministic case, a single observational input is intepreted to
be caused by a range of possible model parameters, described by a probability
distribution.
This statement is true even in the case of a perfect observation of infinite
signal-to-noise, since the underlying model itself comes with uncertainty.
In recent years, there has been a shift towards such methods in several areas
of astronomy, particularly within the exoplanet community.
For example, \citet{wolfgang:2015} considered that the mass-radius
relationship of exoplanets is stochastic, since a particular sized planet
could be have a range of planet masses due to unmodeled variances in
compositions, environment and other complications.
These recent demonstrations in exoplanetary science have prompted us to
consider the need for treating the parent stars in the same probabilistic
framework, with potential applications spanning many fields of astronomy.

The demand for probabilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, it also has a practical purpose.
When using data published in the astronomical literature to, for example, infer
relationships between observed parameters, that inference can be performed as
the final stage in a hierarchical treatment
\citep[see, e.g.][]{foreman-mackey:2014}.
This is useful if, for example, one wants to account for multi-dimensional
uncertainties by marginalizing over the `true' parameter values upon which the
noisy observations are conditioned.
This point is discussed further in \textsection\ref{sec:HBM}.

@@

One of the more recent tools developed to characterize stars is known as
``flicker'' \citep{bastien:2013}.
Flicker is a proxy for the scatter on an 8-hour timescale (denoted as $F_8$)
in a broad visible bandpass time series photometric light curve, such as that
from \textit{Kepler} or the upcoming TESS mission. A more detailed account of
the proceedure to calculate flicker is described in \citet{bastien:2013}. As
shown in \citet{bastien:2013}, flicker displays a remarkable correlation to
the asteroseismically determined parent star surface gravities (\logg).
Turning this around, the observation implies that flicker can be used to
blindly infer surface gravities at the level of $\sim0.1$\,dex, an attractive
proposition given the wealth of photometric light curves available through the
array of exoplanet transit missions flying and scheduled to launch.

@@

\begin{itemize}
\item[{\tiny$\blacksquare$}] Range (defined in \citealt{bastien:2013})
$<1$\,ppt
\item[{\tiny$\blacksquare$}] $4500<T_{\mathrm{eff}}<6500$\,K
\item[{\tiny$\blacksquare$}] $K_P<14$
\item[{\tiny$\blacksquare$}] $1.2 < \log_{10}$($F_8$\,[ppm])$< 2.2$
\end{itemize}

@@

The two models we use to describe the relationships between $F_8$, \logg\ and
\rhostar\ are
\begin{equation}
% 	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
% 	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
    \beta_\rho \log_{10}(\rho_\star), \sigma^2 = \sigma_\rho^2 \right),
\end{equation}
\label{eq:rho}
and
\begin{equation}
% 	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
% 	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
    \log_{10}(g), \sigma^2 = \sigma_g^2 \right).
\end{equation}
\label{eq:logg}
The free parameters of the two models are $\alpha_\rho$, $\beta_\rho$,
$\sigma_\rho$, $\alpha_g$, $\beta_g$ and $\sigma_g$.
These relations are Gaussian distributions with means given by the equation of
a straight line, and standard deviations which describe the intrinsic scatter
about the mean.

We also tested a model in which the additional scatter depends on flicker
itself, defined as
\begin{equation}
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
    \beta_\rho \log_{10}(\rho_\star), \sigma = \sqrt{\sigma_{\rho}^2 +
    \gamma_\rho F_8}\right),
\end{equation}
\label{eq:rho2}
for flicker vs $\rho_\star$ and similarly for \logg.
This model allowed us to determine whether there the magnitude of additional
scatter varied as a function of flicker.
In other words, whether flicker was a better proxy for \logg\ or \rhostar\ for
either dwarf or giant stars.
We found that the maximum {\it a-posteriori} values for the $\gamma$
parameters were consistent with zero: $\gamma_{rho} = 0.006 \pm 0.02, \gamma_g
= -0.01 \pm\ 0.01$, and interpret this as evidence for a constant intrinsic
scatter level across evolutionary stages.

@@

We used a likelihood function which accounts for 2-D uncertainties but does
not allow the intrinsic scatter to be a function of the dependent or
independent variables.
For the relation between flicker and \rhostar, this likelihood function can be
written as
\begin{eqnarray}
	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) =  \\ \nonumber
						      & \exp \left[-\frac{1}{2}
		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
	{\left[\beta \sigma_{F8, n}^2 + (\sigma_{\rho *, n}
	+ \sigma_{\rho})^2\right]}\right]
	\\ \nonumber
	& + \beta^2 \sigma_{F8, n}^2 + (\sigma_{\rho *, n} + \sigma_{\rho})^2,
\end{eqnarray}
\label{eq:likelihood}
and similarly for \logg.
We found that the posterior PDFs for the model parameters obtained using this
likelihood function were consistent with those obtained using a model that
only accounts for the uncertainties on the flicker measurements.
Accounting for uncertainties on $y$ {\it and} $x$ is therefore not essential
in this case but still good practise and will, in general, produce more
accurate model parameters and uncertainties.

@@

We used the uninformative prior for the parameters of a straight line with
unknown variance, outlined in \citet{vanderplas},
\begin{equation}
P(\alpha, \beta, \sigma) \propto \frac{1}{\sigma} \left( 1 + \beta^2
\right)^{-3/2}.
\end{equation}
\label{eq:priors}

@@

\begin{table}
\begin{center}
\caption{Median parameter values with 16th and 84th
	percentile uncertainties.}
\begin{tabular}{lcc}
\hline\hline
Parameter & Median value \\
    \hline
$\alpha_\rho$   &    1.315$\pm 0.008$   \\
$\beta_\rho$    &    -0.526$\pm 0.009$   \\
$\sigma_\rho$   &    0.065$\pm 0.003$   \\
% 1.31476104613 0.00759733599155 0.00767814819025
% -0.525867117083 0.00913323668806 0.00926420817321
% 0.0650160215492 0.00300880167448 0.00288553857806
\hline
$\alpha_g$      &    4.90$\pm 0.05$   \\
$\beta_g$       &   -0.83$\pm 0.01$   \\
$\sigma_g$      &    0.056$\pm 0.003$   \\
% 4.90451472909 0.0514398889837 0.0509386027456
% -0.8252952869 0.0130730719222 0.0131097233196
% 0.0565263079576 0.00279179051697 0.00271277097572
    \hline
\end{tabular}
\end{center}
\end{table}
\label{tab:results}

@@

This is a simple `fitting a line to data' exercise, however it continues the
discussion of probabilistic modelling that is an active topic within the
fields of exoplanet and stellar astronomy.
We used Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
scatter in the relationship between flicker, surface gravity and density and
included the effects of the non-negligable two-dimensional observational
uncertainties.
% Although none of these methods, nor the data used here are new taken alone,
% we use the example of flicker to demonstrate the ideas and the importance of
% probabilistic, hierarchical inference.
Relationships between astronomical parameters are almost always
non-deterministic; an element of stochasticity effects the physical parameters
of stars so one can never perfectly predict $y$ given an observation of $x$.
We advocate a probabilistic approach in both the `fitting the model to data'
step, {\it and} when using an empirically calibrated model to predict
parameter values.
The fitting stage benefits because if the relationships between parameters are
falsely assumed to be deterministic, they will be skewed by data points with
unrepresentative uncertainties.
The prediction stage benefits from the stochastic treatment both because a
probability distribution is in many ways more representative of an observation
than a point estimate, and because posterior PDF samples can be used in
subsequent studies (provided the prior used during the fitting process is
described).

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}
Fox, J., 1997, Applied Regression Analysis, Linear Models, and Related Methods
(Thousand Oaks:Sage Publications, Inc.)
\bibitem[\protect\citeauthoryear{Fuller}{1987}]{fuller:1987}
Fuller, W.~A., 1987, Wiley Series in Probability and Mathematical Statistics
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010b}]{hogg:2010b}
Hogg, D.~W., Bovy, J. \& Lang, D., 2010
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kelly}{2007}]{kelly:2007} Kelly, B.~C., 2007,
ApJ, 665, 1489
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]
{seager:2003}Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
\bibitem[VanderPlas(2014)]{vanderplas}
VanderPlas, J.\ 2014, arXiv:1411.5018
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}

>>>>

An alternative approach for inferring model parameters is to allow
relationships between observables to be stochastic.
% Unlike the deterministic case, a single observational input is interpreted to
% be caused by a range of possible model parameters, described by a probability
% distribution.
% This statement is true even in the case of a perfect observation of infinite
% signal-to-noise, since the underlying model itself comes with uncertainty.
In recent years, there has been a shift towards such methods in several areas
of astronomy, particularly within the exoplanet community.
For example, \citet{wolfgang:2015} considered that the mass-radius
relationship of exoplanets is stochastic, since a particular sized planet
could be have a range of planet masses due to unmodeled variances in
compositions, environment and other complications.
These recent demonstrations in exoplanetary science have prompted us to
consider the need for treating the parent stars in the same probabilistic
framework, with potential applications spanning many fields of astronomy.

The demand for probabilistic stellar parameters is not only motivated by the
fact that probability distributions are far more representative of our
`beliefs' about astrophysical parameters, it also has a practical purpose.
When using data published in the astronomical literature to, for example,
infer relationships between parameters that are themselves the product of an
inference process (for example, exoplanet transit depth and period), inference
can be performed as the final stage in a hierarchical treatment \citep[see,
e.g.][]{foreman-mackey:2014}.
% This is useful if, for example, one wants to account for multi-dimensional
% uncertainties by marginalizing over the `true' parameter values upon which the
% noisy observations are conditioned.
Studies such as these are benefited by posterior PDF samples, rather than
point estimates of inferred properties.
% This point is discussed further in \textsection\ref{sec:HBM}.

@@

One of the more recent tools developed to characterize stars is known as
``flicker'' \citep{bastien:2013}.
Flicker is a proxy for the scatter on an 8-hour timescale (denoted as $F_8$)
in a broad visible bandpass time series photometric light curve, such as that
from \textit{Kepler} or the upcoming TESS mission. A more detailed account of
the proceedure to calculate flicker is described in \citet{bastien:2013}. As
shown in \citet{bastien:2013}, flicker displays a remarkable correlation to
the asteroseismically determined parent star surface gravities (\logg).
Turning this around, the observation implies that flicker can be used to
empirically infer surface gravities at the level of $\sim0.1$\,dex, an
attractive proposition given the wealth of photometric light curves available
through the array of exoplanet transit missions flying and scheduled to
launch.

@@

\begin{itemize}
\item[{\tiny$\blacksquare$}] Range (defined in \citealt{bastien:2013})
$<1000$\,ppm
\item[{\tiny$\blacksquare$}] $4500<T_{\mathrm{eff}}<6500$\,K
\item[{\tiny$\blacksquare$}] $K_P<14$
\item[{\tiny$\blacksquare$}] $1.2 < \log_{10}$($F_8$\,[ppm])$< 2.2$
\end{itemize}

@@

The two models we use to describe the relationships between $F_8$, \logg\ and
\rhostar\ are
\begin{equation}
% 	\log_{10}(\rho_\star) \sim \mathcal{N} \left(\mu = \alpha + \beta
% 	\log_{10}(F_8), \sigma = \sqrt{\sigma_{\rho}^2 + \gamma F_8}\right),
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_\rho^2 \right),
\end{equation}
\label{eq:rho}
and
\begin{equation}
% 	\log_{10}(g) \sim \mathcal{N}\left(\mu = \delta + \epsilon \log_{10}(F_8),
% 	\sigma = \sqrt{\sigma_g^2 + \zeta F_8}\right).
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_g + \beta_g
    \log_{10}(g), \sigma^2 + \sigma_g^2 \right).
\end{equation}
\label{eq:logg}
The free parameters of the two models are $\alpha_\rho$, $\beta_\rho$,
$\sigma_\rho$, $\alpha_g$, $\beta_g$ and $\sigma_g$.
These relations are Gaussian distributions with means given by the equation of
a straight line, and standard deviations which describe the intrinsic scatter
about the mean.

We also tested a model in which the additional scatter depends on flicker
itself, defined as
\begin{equation}
	\log_{10}(F_8) \sim \mathcal{N} \left(\mu = \alpha_\rho +
    \beta_\rho \log_{10}(\rho_\star), \sigma^2 + \sigma_{\rho}^2 +
    \gamma_\rho F_8 \right),
\end{equation}
\label{eq:rho2}
for flicker vs $\rho_\star$ and similarly for \logg.
This model allowed us to determine whether there the magnitude of additional
scatter varied as a function of flicker.
In other words, whether flicker was a better proxy for \logg\ or \rhostar\ for
either dwarf or giant stars.
We found that the maximum {\it a-posteriori} values for the $\gamma$
parameters were consistent with zero: $\gamma_{rho} = 0.006 \pm 0.02, \gamma_g
= -0.01 \pm\ 0.01$, and interpret this as evidence for a constant intrinsic
scatter level across evolutionary stages.

@@

We used a likelihood function which accounts for 2-D uncertainties but does
not allow the intrinsic scatter to be a function of the dependent or
independent variables.
For the relation between flicker and \rhostar, this likelihood function can be
written as
\begin{eqnarray}
	& p(F_8|\rho_\star, \alpha, \beta, \sigma_{\rho}) \propto  \\ \nonumber
						      & \exp \left[-\frac{1}{2}
		\sum_{n=1}^N \frac{[F_{8n}-(\alpha + \beta \rho_{*n})]^2}
	{\left[\beta \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2
	+ \sigma_{\rho}^2\right]}\right]
	\\ \nonumber
	& + \beta^2 \sigma_{F8, n}^2 + \sigma_{\rho *, n}^2 + \sigma_{\rho}^2,
\end{eqnarray}
\label{eq:likelihood}
and similarly for \logg.
We found that the posterior PDFs for the model parameters obtained using this
likelihood function were consistent with those obtained using a model that
only accounts for the uncertainties on the flicker measurements.
Accounting for uncertainties on $y$ {\it and} $x$ is therefore not essential
in this case but still good practise and will, in general, produce more
accurate model parameters and uncertainties.

@@

We used the uninformative prior for the parameters of a straight line for
data with unknown uncertainties, outlined in \citet{vanderplas},
\begin{equation}
P(\alpha, \beta, \sigma) \propto \frac{1}{\sigma} \left( 1 + \beta^2
\right)^{-3/2}.
\end{equation}
\label{eq:priors}

@@

\begin{table}
\begin{center}
\caption{Median parameter values with 1$\sigma$ uncertainties.}
\begin{tabular}{lcc}
\hline\hline
Parameter & Median value \\
    \hline
$\alpha_\rho$   &    1.315$\pm 0.008$   \\
$\beta_\rho$    &    -0.526$\pm 0.009$   \\
$\sigma_\rho$   &    0.065$\pm 0.003$   \\
% 1.31476104613 0.00759733599155 0.00767814819025
% -0.525867117083 0.00913323668806 0.00926420817321
% 0.0650160215492 0.00300880167448 0.00288553857806
\hline
$\alpha_g$      &    4.90$\pm 0.05$   \\
$\beta_g$       &   -0.83$\pm 0.01$   \\
$\sigma_g$      &    0.056$\pm 0.003$   \\
% 4.90451472909 0.0514398889837 0.0509386027456
% -0.8252952869 0.0130730719222 0.0131097233196
% 0.0565263079576 0.00279179051697 0.00271277097572
    \hline
\end{tabular}
\end{center}
\end{table}
\label{tab:results}

@@

This is a simple `fitting a line to data' exercise, however it continues the
discussion of probabilistic modelling that is an active topic within the
fields of exoplanet and stellar astronomy.
We used Hierarchical Bayesian Modelling (HBM) to constrain the intrinsic
scatter in the relationship between flicker, surface gravity and density and
included the effects of the non-negligable two-dimensional observational
uncertainties.
% Although none of these methods, nor the data used here are new taken alone,
% we use the example of flicker to demonstrate the ideas and the importance of
% probabilistic, hierarchical inference.
Relationships between astronomical parameters are almost always
non-deterministic; an element of stochasticity effects the physical parameters
of stars so one can never perfectly predict $y$ given an observation of $x$.
We advocate a probabilistic approach in both the `fitting the model to data'
step, {\it and} when using an empirically calibrated model to predict
parameter values.
The fitting stage benefits because if the relationships between parameters are
falsely assumed to be deterministic, they will be skewed by data points with
uncertainties that only represent measurement error and no additional scatter.
The prediction stage benefits from the stochastic treatment both because a
probability distribution is in many ways more representative of an observation
than a point estimate, and because posterior PDF samples can be used in
subsequent studies (provided the prior used during the fitting process is
described).

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}
Fox, J., 1997, Applied Regression Analysis, Linear Models, and Related Methods
(Thousand Oaks:Sage Publications, Inc.)
\bibitem[\protect\citeauthoryear{Fuller}{1987}]{fuller:1987}
Fuller, W.~A., 1987, Wiley Series in Probability and Mathematical Statistics
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010b}]{hogg:2010b}
Hogg, D.~W., Bovy, J. \& Lang, D., 2010
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kelly}{2007}]{kelly:2007} Kelly, B.~C., 2007,
ApJ, 665, 1489
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]
{seager:2003}Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
\bibitem[VanderPlas(2014)]{vanderplas}
VanderPlas, J.\ 2014, arXiv:1411.5018
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}
